---
title: "eplPredictor"
output: html_document
date: "2025-05-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of 2025 Premier League Results

## Reading in and preparing the dataset
```{r}
# preparing the dataset
library(readr)
# read in the epl_merged csv file
epl_merged <- read_csv("C:/Users/rache/Documents/CodingProjects/SoccerPredictor/eplMerged.csv")
View(epl_merged)
attach(epl_merged)
# print out the names of the columns in the epl_merged dataset
names(epl_merged)
# get the dimensions of the dataset
dim(epl_merged)
```

## Exploratory Data Analysis
Response variable (Y): Wins ("W")

Potential Predictor variables (X): 
-Goal Difference ("GD")
-Attendance (per game)
-Number of players ("num_players")
-Age (average age for a team)
-Possession ("Poss")
-Goals + Assists ("G_A")
-Yellow Cards ("CrdY")
-Red Cards ("CrdR")
-Progressive Carries ("PrgC")
-Progressive passes ("PrgP")
-Total Shots ("Sh")
-Shots on Target ("SoT")
-Goals per shot ("G_per_Sh")
-Average Shot Distance ("Dist")
-Total Pass Completion Percentage ("Total_Cmp")
-Short Pass Completion Percentage ("Short_Cmp")
-Medium Pass Completion Percentage ("Medium_Cmp")
-Long Pass Completion Percentage ("Long_Cmp")
-Successful Take-ons that lead to a shot attempt ("TO")
-Fouls drawn that lead to a shot attempt ("Fld")


### Understanding the Dataset - Summary Statistics
```{r}
# generating summary statistics for each column
summary(epl_merged)
```

### Understanding the DataSet - Graphs
```{r}
# Printing out relationship between Wins vs Goal Difference, Attendance, # of players, Age
predictors1_4 <- c("GD", "Attendance", "num_players", "Age")
par(mfrow=c(2,2))
for (var in predictors1_4) {
    plot(epl_merged[[var]], epl_merged$W, xlab=var, ylab="Wins", main=paste("w vs", var))
    abline(lm(W~epl_merged[[var]], data=epl_merged))
}

# Printing out relationship between Wins vs Possession, Goals + Assists, Yellow Cards, Red Cards
predictors5_8 <- c("Poss", "G_A", "CrdY", "CrdR")
for (var in predictors5_8) {
    plot(epl_merged[[var]], epl_merged$W, xlab=var, ylab="Wins", main=paste("w vs", var))
    abline(lm(W~epl_merged[[var]], data=epl_merged))
}

# Printing out relationship between Wins vs Progressive Carries, Progressive Passes, Total Shots, Shots on Target
predictors9_12 <- c("PrgC", "PrgP", "Sh", "SoT")
for (var in predictors9_12) {
    plot(epl_merged[[var]], epl_merged$W, xlab=var, ylab="Wins", main=paste("w vs", var))
    abline(lm(W~epl_merged[[var]], data=epl_merged))
}


# Printing out relationship between Wins vs Goals per shot, Average Shot Distance, Total Pass Completion %, Short Pass Completion %
predictors13_16 <- c("G_per_Sh","Dist", "Total_Cmp", "Short_Cmp")
for (var in predictors13_16) {
    plot(epl_merged[[var]], epl_merged$W, xlab=var, ylab="Wins", main=paste("w vs", var))
    abline(lm(W~epl_merged[[var]], data=epl_merged))
}

# Printing out relationship between Wins vs Medium Pass Completion %, Long Pass Completion %, Successful Take-ons that lead to a shot attempt, Fouls Drawn that lead to a shot attempt
predictors17_20 <- c("Medium_Cmp", "Long_Cmp", "TO", "Fld")
for (var in predictors17_20) {
    plot(epl_merged[[var]], epl_merged$W, xlab=var, ylab="Wins", main=paste("w vs", var))
    abline(lm(W~epl_merged[[var]], data=epl_merged))
}

```

## Initial Modeling
```{r}
# getting all predictors to include in the initial model
predictors <- c("GD", "Attendance", "num_players", "Age", "Poss", "G_A", "CrdY", "CrdR", "PrgC", "PrgP", "Sh", "SoT", "G_per_Sh", "Dist", "Total_Cmp", "Short_Cmp", "Medium_Cmp", "Long_Cmp", "TO", "Fld")

```

